Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-01-06 paper/paper.tex
pass

In the following, we illustrate how \tname{} proceeds for obtaining
the rule for the SYN Flood attack. One challenge in synthesizing
rules is to discover the set of options in a rule that makes the \nids\ to
capture the malicious traffic, but allow the benign traffic to go through.  The
other parts of the rule (\ie{}, the action and the header) are either
inferred or left constant. \tname{} sets the action header to
\CodeIn{alert} and sets the header to \CodeIn{<proto> any any -> any
  any}, indicating that it infers the target protocol and uses the
most general flow description possible, \ie{}, it instructs the
\nids\ to analyze the traffic flowing from/to any address and port.
We understand that this is domain-specific information and conjecture
that the system administrator would know how to customize.

\tname{} proceeds as follows to synthesize rules for this
attack. First, it analyzes the positive traffic and extracts options
expressed in that traffic. For this example, it finds the following
set of options:

\begin{figure}[h]
  \vspace{-3ex}
  \lstinputlisting[language=C,numbers=none,frame=none,keywords={flags,threshold}]{synflood.suricata.synth}
%  \caption{Suricata rule for SYN Flood Attacks.}
  %  \label{fig:synflood-example-synt}
  \vspace{-3ex}  
\end{figure}

\Mar{<~ Lucas/Guilheme, logo acima a gente falou que threshold nao
  poderia ser inferido a partir de uma unica mensagem e aqui a gente
  fala isto. Vcs. podem explicar o que realmente acontece?}

Note that, by construction, a rule that uses this set of options
captures the positive traffic as these options originated from the
positive traffic. Unfortunately, this rule is overspecified. It would
unlikely capture different manifestations of the same
attack. Informally, the rule includes ``too many'' characteristics of
the malicious traffic used to define it. To address this overfitting
issue, \tname{} minimizes the rule preserving the invariant that
plausible rules should only capture the positive traffic. \tname{}
potentially finds several minimal rules. The list below shows the
options for the five shortest rules it produces for this attack.

\Mar{<~ Lucas/Guilheme, vcs. podem atualizar esta informacao?}

>>>>

In the following, we illustrate how \tname{} proceeds for obtaining
the rule for the SYN Flood attack. The challenge in synthesizing rules
is to discover the set of options in a rule that makes the \nids\ to
capture the malicious traffic, but allow the benign traffic to go
through.  The other parts of the rule (\ie{}, the action and the
header) are either inferred or left constant. For example, in this
case, \tname{} sets the action header to \CodeIn{alert} and sets the
header to \CodeIn{<proto> any any -> any any}, indicating that it
infers the target protocol and uses the most general flow description
possible, \ie{}, it instructs the \nids\ to analyze the traffic
flowing from/to any address and port.  We understand that this is
domain-specific information and assume that the system administrator
would know how to customize it.

\tname{} proceeds as follows to synthesize rules for this
attack. First, it analyzes the positive (\ie, malicious) traffic and
extracts options expressed in that traffic. In this case, the
malicious traffic is characterized as a sequence of 20
messages. \tname{} finds the following set of options at this stage:

\begin{figure}[h]
  \vspace{-3ex}
  \lstinputlisting[language=C,numbers=none,frame=none,keywords={flags,threshold,window}]{synflood.suricata.synth}
%  \caption{Suricata rule for SYN Flood Attacks.}
  %  \label{fig:synflood-example-synt}
  \vspace{-3ex}  
\end{figure}

%% \Mar{<~ Lucas/Guilheme, logo acima a gente falou que threshold nao
%%   poderia ser inferido a partir de uma unica mensagem e aqui a gente
%%   fala isto. Vcs. podem explicar o que realmente acontece?}
%% \Luc{O input considerado para ataques de flood eh um conjunto de pacotes e não apenas um pacote unico, nesse caso um conjunto com 20 pacotes. A opção de threshold é incluida na regra quando o input contem mais de um pacote, o parametro count recebe o valor da quantidade de pacotes do conjunto de entrada.}

\tname{} only reports a rule if the \nids\ (\ie{}, Suricata)
instantiated with that rule is able to capture the traffic passed on
input. Consequently, by construction, a rule that uses this set of
options captures the malicious traffic. Unfortunately, this rule is
potentially overspecified. Informally, the rule includes ``too many''
characteristics of the malicious traffic used to define it and could
potentially result in false negatives.

Second, \tname{} uses negative (\ie{}, benign) traffic to address the
overfitting issue. \tname{} searches for alternative rules that
preserves the invariant that a rule should only capture the positive
traffic. It uses the rule obtained in the previous step to bootstrap
the search.

Finally, \tname{} ranks the inferred rules according to heuristics
extracted from a database of existing \tname{} rules. One example
heuristic is \Fix{...}. Instead of traffic, this step looks for the
structure of existing rules. The list below shows the rule options for
the five shortest rules \tname{} produces for this attack.

\Mar{<~ Lucas/Guilheme, há algo o que remover? talvez ``window:64''?
  Qual a correspondencia de ``flags:S,12'' com ``flags:S''? Por favor,
  respondam isto logo.}  \Gui{}
